Kindly Bent to Free Us

### 2018-11-08 kindly/infer/infer.tex
Fix/Extend the type inference rules.

\subsection{Constraint language}

@@

\paragraph{Solved forms}

@@

@@

We note $\inferK{(\unif,C)}{\E}{\tau}{k}$ the fact that type $\tau$ has kind $k$
in environment $\E$ with constraints $C$ and unifier $\unif$. From an
algorithmic point of view, $\E$ and $\tau$ are the input parameters of
our inference procedure.

@@

\begin{figure*}[h]
  \input{infer/validity}
  \caption{Type validity rules -- $\tyval{C}{\E}{\tau}$ }
  \input{infer/rules}
  \caption{Inference rules -- $\inferW{\Sigma}{(C,\psi)}{\bf{\E}}{\bf{e}}{\tau}$ }
\end{figure*}

\section*{Todo}

\begin{itemize}
\item Add type constructors
\item Properly ensure that it respects HM(X) (cylindric, \dots)
\item Show principal type inference:
  \begin{itemize}
  \item Principal constraint system
  \item Regular constraint system: $\Ceq{\tau}{\tau'} \implies \fv{\tau} = \fv{\tau'}$.
  \item Solved forms are in simplified form.
    $C\in S, \entail{C}{\Ceq{\tau}{\tau'}} \implies \entail{}{\Ceq{\tau}{\tau'}}$.
  \end{itemize}
\item Show equivalence with the logic-based system.

>>>>

\subsection{Preliminaries}

\subsubsection{Constraint language}

@@

\subsubsection{Solved forms}

@@

\subsubsection{Usage maps}

\TODO{Explain $\Sigma$}

@@

We note $\inferK{(C,\unif)}{\bf{\E}}{\bf{\tau}}{k}$ when type $\tau$ has kind $k$
in environment $\E$ under constraints $C$ and unifier $\unif$. From an
algorithmic point of view, $\E$ and $\tau$ are the input parameters of
our inference procedure.
We present the kind inference algorithm as a set of syntax-directed rules in
\cref{rules:kinding}.
Since higher-kinded types are not supported, the type application
rule and the type constructor rule are merged in {\sc KApp}.
Additionally, type variables must be of a simple kind in rule {\sc KVar}.
Note that in the case of a type constructor with no argument, the {\sc KApp}
rule degenerates to a simpler form which is similar to the {\sc KVar} rule.
Kind schemes are instantiated in the {\sc KApp} and {\sc KVar} rules by creating
fresh kind variables and the associated substitution.

\TODO{Explain more?}

\begin{figure}[h]
  \centering
  \input{infer/kinds}
  \caption{Kind inference rules -- $\inferK{(C,\unif)}{\E}{\tau}{k}$}
  \label{rules:kinding}
\end{figure}

@@

We reformulate the HM(X) type inference in the context of our affine type
system. The main difference compared to HM(X) are noted in \addlin{blue}.
We note $\inferW{\addlin{\Sigma}}{(C,\unif)}{\bf{\E}}{\bf{e}}{\tau}$ when
$e$\ as type $\tau$ in $\E$ under the constraints $C$ and unifier $\unif$.
$\Sigma$ is a map which associates free variables in $e$ to
their kinds.
As before, $\E$ and $e$ are the input parameters of the inference
algorithm. The syntax-directed rules are shown in \cref{rules:typing}.

\begin{figure*}[h]
  \input{infer/typing}
  \caption{Type Inference rules -- $\inferW{\Sigma}{(C,\psi)}{\bf{\E}}{\bf{e}}{\tau}$ }
  \label{rules:typing}
\end{figure*}

\subsection{Constraint normalization and generalization}

\TODO{}

\subsection{Principality}

\TODO{}

\begin{itemize}
\item Properly ensure that it respects HM(X) (cylindric, \dots)
\item Show principal type inference:
  \begin{itemize}
  \item Principal constraint system
  \item Regular constraint system: $\Ceq{\tau}{\tau'} \implies \fv{\tau} = \fv{\tau'}$.
  \item Solved forms are in simplified form.
    $C\in S, \entail{C}{\Ceq{\tau}{\tau'}} \implies \entail{}{\Ceq{\tau}{\tau'}}$.
  \end{itemize}
\item Show equivalence with the logic-based system.

### 2018-11-08 kindly/lang.tex
Fix/Extend the type inference rules.

The syntax of \lang is presented in \cref{grammar} and follows
traditional ML languages with let bindings.
Data constructors are introduced by $\app{C}{x}$ and eliminated
with by $\matchin{C}{x}{e}{e'}$.
Types are composed of type variable $\tvar$, parameterized type constructors
$\T{t}$ and arrow types annotated with a kind, noted $\tau \tarr{k} \tau$.
Kinds can be either a kind variable $\kvar$, affine $\kaff$ or unrestricted $kun$.
Type schemes and kind schemes are guarded by a set of constraints.

Environments are composed of value and type bindings, along with type
declarations. Type declarations, of the form
$\tydecl{t}{\kschm}{C}{\tau}$, introduce both a new type constructor $\T{t}$ and
a data constructor $C$.

>>>>

The syntax of \lang is presented in \cref{grammar} and follows
traditional ML languages with let bindings.
Data constructors are introduced by $\introK{K}{e}$ and eliminated
with by $\elimK{K}{e}$.
Types are composed of type variable $\tvar$, parameterized type constructors
$\T{t}$ and arrow types annotated with a kind, noted $\tau \tarr{k} \tau$.
Kinds can be either a kind variable $\kvar$, affine $\kaff$ or unrestricted $kun$.
Type schemes and kind schemes are guarded by a set of constraints.

Environments are composed of value and type bindings, along with type
declarations. Type declarations, of the form
$\tydecl{t}{\kschm}{K}{\tau}$, introduce both a new type constructor $\T{t}$ and
a data constructor $K$.
